---
title: "K estimation"
author: "MKAUFMAN"
date: "`r Sys.Date()`"
output: html_document
---
##Matthew Kaufman
#matthew.kaufman@pnnl.gov 
#Pacific Nortwest National Laboratory River Corridor SFA
#This script takes measured and modeled hydraulic conditions (velocity, slope, depth, and discharge) for various stream reaches and applies an empirical formula to estimate K600 from them

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## import depth and width data, slope data, discharge and velocity data, and combine
```{r message=FALSE}
data1<-read.csv("SSS_Water_Depth_Summary.csv",skip=8) #from this data package: https://data.ess-dive.lbl.gov/datasets/doi:10.15485/1969566

data2<-read.csv("SSSflow_velocity.csv") #We obtained the streamflow and flow velocity from NOAA National Water Model (NWM) retrospective dataset using version 2.1 of NWM. The dataset provides a 42-year (February through December 2020) hourly simulation at all sampling locations (except COMID 24125857). We computed the monthly averaged streamflow and velocity in August across these simulation years.

colnames(data2)<-c("Site_ID","comid","discharge","velocity")

data3<-read.csv("SSS_Slope.csv") #from NHDplus v2.1

colnames(data3)<-c("Site_ID","slope")
data1$Average_Depth_m=data1$Average_Depth/100

data1 <-data1[order(data1$Site_ID),]
data2 <-data2[order(data2$Site_ID),]
data3 <-data3[order(data3$Site_ID),]
data = cbind.data.frame(data1$Site_ID,data1$Average_Depth_m,data2$discharge,data2$velocity,data3$slope)
colnames(data)<-c("Site_ID","Average_Depth_m","discharge","velocity","slope")
```


## Estamate k600 using Raymond 2012 table 2 equation 7:  k600=4725 · (VS)^0.86 · Q^-0.14 · D^0.66
#k600 [m/d], velocity V [m/s], slpe S [-], depth D [m], discharge Q [m^3/s]

```{r message=FALSE}
data$k600_md<-4725*(data$velocity*data$slope)^0.86*data$discharge^-0.14*data$Average_Depth_m^0.66

output<-data.frame(data[c("Site_ID","k600_md")])
colnames(output)=c("Site_ID","K600")
output$K600<-round(output$K600,2)

header <- tibble(header = c('# HeaderRows_4',
                            '# HeaderRows_Format: Column_Header; Unit; InstallationMethod_ID; Instrument_Summary', 
                            'k600; meters_per_day; N/A; Calculated using (1) depth (2) flow (3) velocity and (4) slope as inputs for the Raymond et al. (2012) table 2 equation 7.'))

write_csv(header, 'SSS_k600.csv', col_names = F)

write_csv(output, 'SSS_k600.csv', append = T, col_names = T)

```